---
title: "Deconvolution"
author: "Dongyue Xie"
date: "2019-12-10"
output:
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE)
```

## Introduction

check bulk data library size: vary from few hundred(100-200) to few thousands(1600). Here 

```{r}
GSE50244bulkeset <- readRDS("~/misc/data/deconv/GSE50244bulkeset.rds")
GSE50244bulkeset
summary(as.numeric(apply(exprs(GSE50244bulkeset),2,mean)))

Mousebulkeset <- readRDS("~/misc/data/deconv/Mousebulkeset.rds")
summary(as.numeric(apply(exprs(Mousebulkeset),2,mean)))
```

We have scRNA-seq data from Segerstolpe et al. (2016) including the 1097 cells from 6 healthy subjects, taken from [here](https://xuranw.github.io/MuSiC/articles/pages/data.html). This dataset is non-UMI so the read counts are huge. In the following simulation study, 4 cell types -  acinar, alpha, beta, and ductal cells are included. 

```{r}
EMTAB.eset = readRDS('data/deconv/EMTABesethealthy.rds')
EMTAB.eset

table(EMTAB.eset$cellType)
```

Now obtain the true gene relative expression in each cell type.

Preprocess the data: 1. remove genes appearing in less than 10 cells; 2. remove top $5\%$ expressed genes  

```{r}
cell_types = c('acinar','alpha','beta','ductal')

###remove genes appeared in too few cells
###remove genes that are overly expressed
#rm.idx = which(rowSums((exprs(EMTAB.eset)[,which(EMTAB.eset$cellType%in%cell_types)])!=0)<10)
#rr = rowSums((exprs(EMTAB.eset)[,which(EMTAB.eset$cellType%in%cell_types)]))
###rm.idx = which(rr==0)
###rm.idx1 = which(rr<=quantile(rr[-rm.idx],0.01))
#rm.idx2 = which(rr>=quantile(rr[-rm.idx],0.99))
#rm.idx = unique(c(rm.idx,rm.idx2))

acinar = exprs(EMTAB.eset)[,which(EMTAB.eset$cellType=='acinar')]
alpha = exprs(EMTAB.eset)[,which(EMTAB.eset$cellType=='alpha')]
beta = exprs(EMTAB.eset)[,which(EMTAB.eset$cellType=='beta')]
ductal = exprs(EMTAB.eset)[,which(EMTAB.eset$cellType=='ductal')]

rm.idx1 = which(((rowSums(acinar!=0)<=10)|(rowSums(acinar)>=quantile(rowSums(acinar),0.95))))
rm.idx2 = which(((rowSums(alpha!=0)<=10)|(rowSums(alpha)>=quantile(rowSums(alpha),0.95))))
rm.idx3 = which(((rowSums(beta!=0)<=10)|(rowSums(beta)>=quantile(rowSums(beta),0.95))))
rm.idx4 = which(((rowSums(ductal!=0)<=10)|(rowSums(ductal)>=quantile(rowSums(ductal),0.95))))

rm.idx = unique(c(rm.idx1,rm.idx2,rm.idx3,rm.idx4))

acinar = acinar[-rm.idx,]
alpha = alpha[-rm.idx,]
beta = beta[-rm.idx,]
ductal = ductal[-rm.idx,]

```

Check cell library size: cell library sizes are big 

```{r}
summary(colSums(acinar))
summary(colSums(alpha))
summary(colSums(beta))
summary(colSums(ductal))
```


```{r}

# cell type specific gene relative expression
Theta = cbind(rowSums(acinar)/sum(acinar),
              rowSums(alpha)/sum(alpha),
              rowSums(beta)/sum(beta),
              rowSums(ductal)/sum(ductal))

cor(Theta)

kappa(t(Theta)%*%Theta)

# cell library size
S = c(sum(acinar)/ncol(acinar),
      sum(alpha)/ncol(alpha),
      sum(beta)/ncol(beta),
      sum(ductal)/ncol(ductal))

#S=S/100 
```


Generate reference data: 1. obtain $\tilde{\theta}_{gk}$ by summing up read counts and normalize; 2. obtain $y_{k}^{r,+}$ by multiple cell library size and number of cells(set to 100); 3. Generate reference data using the Poisson model

Generate bulk data: 1. simulate library size $y^{b,+}$ from Poisson$(300G)$; 2. generate $x^b_g$ by multiplying $N=400$, $\beta = (0.1,0.2,0.3,0.4)$, and $S$ cell library size then normalize to $\theta^b_g$. 3. generate bulk data using Poisson distribution.

```{r}
G = nrow(Theta)
K = ncol(Theta)

set.seed(12345)
# bulk data library size: 300*number of genes.
# bulk_ls = rpois(1,300*nrow(Theta))
# total number of cells in bulk data
bulk_ncell = 400
# cell proportions
bulk_beta = c(1,2,3,4)
bulk_beta = bulk_beta/sum(bulk_beta)
bulk_p = (bulk_beta * (S))/sum(bulk_beta * (S))
# bulk data gene relative expression.
bulk_X = bulk_ncell*Theta%*%diag(S)%*%bulk_beta
bulk_theta = bulk_X/sum(bulk_X)

ref_ncell = 100
ref_X = Theta%*%diag(S)*ref_ncell

w = rep(1,G)


p_tilde_est = c()
p_tilde_var = c()
p_est = c()
p_var = c()

nrep = 100
for(rep in 1:nrep){
  if(rep%%200==0){print(rep)}
  bulk_ls = rpois(1,300*nrow(Theta))
  Z = matrix(rpois(G*K,ref_X),ncol=K)
  y = rpois(G,bulk_ls*bulk_theta)

  #W = diag(w)
  #beta_hat = solve(t(Z)%*%W%*%Z - diag(colSums(W%*%Z)))%*%t(Z)%*%W%*%y

  U = diag(colSums(Z))
  Zu = Z%*%(solve(U))
  Zuu = Zu%*%solve(U)
  p_tilde_hat = solve(t(Zu)%*%Zu - diag(colSums(Zuu)))%*%t(Zu)%*%y
  p_tilde_est = rbind(p_tilde_est,c(p_tilde_hat))
  #beta_hat = solve(t(Z)%*%Z)%*%t(Z)%*%y

  #beta_hat
  #beta_hat/sum(beta_hat)

  Q = 0
  Sigma=0
  deltas = c()
  for(i in 1:length(y)){ 
    ag = Zu[i,]%*%t(Zu[i,])-diag(Zuu[i,])
    Q = Q + ag
    Delta = (ag%*%p_tilde_hat-y[i]*Zu[i,])
    Sigma = Sigma + Delta%*%t(Delta)
  }
  Q = Q/G
  Sigma = Sigma/G

  J = matrix(nrow = K,ncol=K)
  for(i in 1:K){
    for(j in 1:K){
      if(i==j){
        J[i,j] = sum(p_tilde_hat)-p_tilde_hat[i]
      }else{
        J[i,j] = -p_tilde_hat[i]
      }
    }
  }

  J = J/sum(p_tilde_hat)^2

  asyV = ((J)%*%solve(Q)%*%Sigma%*%solve(Q)%*%t(J))/G

  p_hat = p_tilde_hat/sum(p_tilde_hat)
  p_est = rbind(p_est,c(p_hat))
  
  p_tilde_var = rbind(p_tilde_var,diag(solve(Q)%*%Sigma%*%solve(Q)))
  
  p_var = rbind(p_var,diag(asyV))

}


#### coverage before delta method:

#ci_l = p_tilde_est - qnorm(0.975)*sqrt(p_tilde_var)
#ci_r = p_tilde_est + qnorm(0.975)*sqrt(p_tilde_var)

#coverage

#coverage = ((rep(1,100)%*%t(bulk_p))>=ci_l) & ((rep(1,100)%*%t(bulk_p))<=ci_r)
#apply(coverage,2,mean)

#### converage after delta method

ci_l = p_est - qnorm(0.975)*sqrt(p_var)
ci_r = p_est + qnorm(0.975)*sqrt(p_var)

#coverage

coverage = ((rep(1,100)%*%t(bulk_p))>=ci_l) & ((rep(1,100)%*%t(bulk_p))<=ci_r)
apply(coverage,2,mean)

#for(i in 1:4){
#  plot(ci_l[,i],type='l',ylim=range(c(ci_l[,i],ci_r[,i])))
#  lines(ci_r[,i],col='grey80')
#  lines(rep(bulk_p[i],100),lty = 2)
#}

```

how about decrease bulk library size to $50G$?

```{r}
G = nrow(Theta)
K = ncol(Theta)

set.seed(12345)
# bulk data library size: 300*number of genes.
# bulk_ls = rpois(1,300*nrow(Theta))
# total number of cells in bulk data
bulk_ncell = 400
# cell proportions
bulk_beta = c(1,1,2,16)
bulk_beta = bulk_beta/sum(bulk_beta)
bulk_p = (bulk_beta * (S))/sum(bulk_beta * (S))
# bulk data gene relative expression.
bulk_X = bulk_ncell*Theta%*%diag(S)%*%bulk_beta
bulk_theta = bulk_X/sum(bulk_X)

ref_ncell = 100
ref_X = Theta%*%diag(S)*ref_ncell

w = rep(1,G)


p_tilde_est = c()
p_tilde_var = c()
p_est = c()
p_var = c()

nrep = 100
for(rep in 1:nrep){
  if(rep%%10==0){print(rep)}
  #bulk_ls = rpois(1,50*nrow(Theta))
  bulk_ls = 50*nrow(Theta)
  Z = matrix(rpois(G*K,ref_X),ncol=K)
  y = rpois(G,bulk_ls*bulk_theta)

  #W = diag(w)
  #beta_hat = solve(t(Z)%*%W%*%Z - diag(colSums(W%*%Z)))%*%t(Z)%*%W%*%y

  U = diag(colSums(Z))
  Zu = Z%*%(solve(U))
  Zuu = Zu%*%solve(U)
  p_tilde_hat = solve(t(Zu)%*%Zu - diag(colSums(Zuu)))%*%t(Zu)%*%y
  p_tilde_est = rbind(p_tilde_est,c(p_tilde_hat))
  #beta_hat = solve(t(Z)%*%Z)%*%t(Z)%*%y

  #beta_hat
  #beta_hat/sum(beta_hat)

  Q = 0
  Sigma=0
  deltas = c()
  for(i in 1:length(y)){ 
    ag = Zu[i,]%*%t(Zu[i,])-diag(Zuu[i,])
    Q = Q + ag
    Delta = (ag%*%p_tilde_hat-y[i]*Zu[i,])
    Sigma = Sigma + Delta%*%t(Delta)
  }
  Q = Q/G
  Sigma = Sigma/G

  J = matrix(nrow = K,ncol=K)
  for(i in 1:K){
    for(j in 1:K){
      if(i==j){
        J[i,j] = sum(p_tilde_hat)-p_tilde_hat[i]
        #J[i,j] = sum(p_tilde_hat)
      }else{
        J[i,j] = -p_tilde_hat[i]
        #J[i,j] = 0
      }
    }
  }

  J = J/sum(p_tilde_hat)^2

  asyV = ((J)%*%solve(Q)%*%Sigma%*%solve(Q)%*%t(J))/G

  p_hat = p_tilde_hat/sum(p_tilde_hat)
  p_est = rbind(p_est,c(p_hat))
  
  p_tilde_var = rbind(p_tilde_var,diag(solve(Q)%*%Sigma%*%solve(Q))/G)
  
  p_var = rbind(p_var,diag(asyV))

}

ci_l = p_est - qnorm(0.975)*sqrt(p_var)*4
ci_r = p_est + qnorm(0.975)*sqrt(p_var)*4

#coverage

coverage = ((rep(1,100)%*%t(bulk_p))>=ci_l) & ((rep(1,100)%*%t(bulk_p))<=ci_r)
apply(coverage,2,mean)

#for(i in 1:4){
#  plot(ci_l[,i],type='l',ylim=range(c(ci_l[,i],ci_r[,i])))
#  lines(ci_r[,i],col='grey80')
#  lines(rep(bulk_p[i],100),lty = 2)
#}

```


how about change $beta$ to be $(0.4,0.3,0.2,0.1)$?

```{r}
G = nrow(Theta)
K = ncol(Theta)

set.seed(12345)
# bulk data library size: 300*number of genes.
# bulk_ls = rpois(1,300*nrow(Theta))
# total number of cells in bulk data
bulk_ncell = 400
# cell proportions
bulk_beta = c(4,3,2,1)
bulk_beta = bulk_beta/sum(bulk_beta)
bulk_p = (bulk_beta * (S))/sum(bulk_beta * (S))
# bulk data gene relative expression.
bulk_X = bulk_ncell*Theta%*%diag(S)%*%bulk_beta
bulk_theta = bulk_X/sum(bulk_X)

ref_ncell = 100
ref_X = Theta%*%diag(S)*ref_ncell

w = rep(1,G)


p_tilde_est = c()
p_tilde_var = c()
p_est = c()
p_var = c()

nrep = 100
for(rep in 1:nrep){
  if(rep%%200==0){print(rep)}
  bulk_ls = rpois(1,300*nrow(Theta))
  Z = matrix(rpois(G*K,ref_X),ncol=K)
  y = rpois(G,bulk_ls*bulk_theta)

  #W = diag(w)
  #beta_hat = solve(t(Z)%*%W%*%Z - diag(colSums(W%*%Z)))%*%t(Z)%*%W%*%y

  U = diag(colSums(Z))
  Zu = Z%*%(solve(U))
  Zuu = Zu%*%solve(U)
  p_tilde_hat = solve(t(Zu)%*%Zu - diag(colSums(Zuu)))%*%t(Zu)%*%y
  p_tilde_est = rbind(p_tilde_est,c(p_tilde_hat))
  #beta_hat = solve(t(Z)%*%Z)%*%t(Z)%*%y

  #beta_hat
  #beta_hat/sum(beta_hat)

  Q = 0
  Sigma=0
  deltas = c()
  for(i in 1:length(y)){ 
    ag = Zu[i,]%*%t(Zu[i,])-diag(Zuu[i,])
    Q = Q + ag
    Delta = (ag%*%p_tilde_hat-y[i]*Zu[i,])
    Sigma = Sigma + Delta%*%t(Delta)
  }
  Q = Q/G
  Sigma = Sigma/G

  J = matrix(nrow = K,ncol=K)
  for(i in 1:K){
    for(j in 1:K){
      if(i==j){
        J[i,j] = sum(p_tilde_hat)-p_tilde_hat[i]
      }else{
        J[i,j] = -p_tilde_hat[i]
      }
    }
  }

  J = J/sum(p_tilde_hat)^2

  asyV = (t(J)%*%solve(Q)%*%Sigma%*%solve(Q)%*%t(J))/G

  p_hat = p_tilde_hat/sum(p_tilde_hat)
  p_est = rbind(p_est,c(p_hat))
  
  p_tilde_var = rbind(p_tilde_var,diag(solve(Q)%*%Sigma%*%solve(Q)))
  
  p_var = rbind(p_var,diag(asyV))

}

ci_l = p_est - qnorm(0.975)*sqrt(p_var)
ci_r = p_est + qnorm(0.975)*sqrt(p_var)

#coverage

coverage = ((rep(1,100)%*%t(bulk_p))>=ci_l) & ((rep(1,100)%*%t(bulk_p))<=ci_r)
apply(coverage,2,mean)

#for(i in 1:4){
#  plot(ci_l[,i],type='l',ylim=range(c(ci_l[,i],ci_r[,i])))
#  lines(ci_r[,i],col='grey80')
#  lines(rep(bulk_p[i],100),lty = 2)
#}

```

The bigger the beta, the lower the coverage??

## pbmc DATA


```{r}
load("~/misc/data/scde/scCD14.RData")
load("~/misc/data/scde/scCD4.RData")
load("~/misc/data/scde/scCD8.RData")
#load("~/misc/data/scde/scCDT.RData")
load("~/misc/data/scde/scMB.RData")

# remove genes 
CD14 = as.matrix(CD14)
CD4 = as.matrix(CD4)
CD8 = as.matrix(CD8)
#CDT = as.matrix(CDT)
MB = as.matrix(MB)
rm.idx1 = which(((rowSums(CD14!=0)<=10)|(rowSums(CD14)>=quantile(rowSums(CD14),0.95))))
rm.idx2 = which(((rowSums(CD4!=0)<=10)|(rowSums(CD4)>=quantile(rowSums(CD4),0.95))))
rm.idx3 = which(((rowSums(CD8!=0)<=10)|(rowSums(CD8)>=quantile(rowSums(CD8),0.95))))
#rm.idx4 = which(((rowSums(CDT!=0)<=10)|(rowSums(CDT)>=quantile(rowSums(CDT),0.95))))
rm.idx5 = which(((rowSums(MB!=0)<=10)|(rowSums(MB)>=quantile(rowSums(MB),0.95))))

rm.idx = unique(c(rm.idx1,rm.idx2,rm.idx3,rm.idx5))

CD14 = CD14[-rm.idx,]
CD4 = CD4[-rm.idx,]
CD8 = CD8[-rm.idx,]
#CDT = CDT[-rm.idx,]
MB = MB[-rm.idx,]

Theta = cbind(rowSums(CD14)/sum(CD14),
              rowSums(CD4)/sum(CD4),
              rowSums(CD8)/sum(CD8),
              #rowSums(CDT)/sum(CDT),
              rowSums(MB)/sum(MB))

cor(Theta)

kappa(t(Theta)%*%Theta)

# cell library size
S = c(sum(CD14)/ncol(CD14),
      sum(CD4)/ncol(CD4),
      sum(CD8)/ncol(CD8),
      sum(MB)/ncol(MB))
```


```{r}
G = nrow(Theta)
K = ncol(Theta)

set.seed(12345)
# bulk data library size: 300*number of genes.
# bulk_ls = rpois(1,300*nrow(Theta))
# total number of cells in bulk data
bulk_ncell = 400
# cell proportions
bulk_beta = c(1,2,3,4)
bulk_beta = bulk_beta/sum(bulk_beta)
bulk_p = (bulk_beta * (S))/sum(bulk_beta * (S))
# bulk data gene relative expression.
bulk_X = bulk_ncell*Theta%*%diag(S)%*%bulk_beta
bulk_theta = bulk_X/sum(bulk_X)

ref_ncell = 100
ref_X = Theta%*%diag(S)*ref_ncell

w = rep(1,G)


p_tilde_est = c()
p_tilde_var = c()
p_est = c()
p_var = c()

nrep = 100
for(rep in 1:nrep){
  if(rep%%200==0){print(rep)}
  bulk_ls = rpois(1,300*nrow(Theta))
  Z = matrix(rpois(G*K,ref_X),ncol=K)
  y = rpois(G,bulk_ls*bulk_theta)

  #W = diag(w)
  #beta_hat = solve(t(Z)%*%W%*%Z - diag(colSums(W%*%Z)))%*%t(Z)%*%W%*%y

  U = diag(colSums(Z))
  Zu = Z%*%(solve(U))
  Zuu = Zu%*%solve(U)
  p_tilde_hat = solve(t(Zu)%*%Zu - diag(colSums(Zuu)))%*%t(Zu)%*%y
  p_tilde_est = rbind(p_tilde_est,c(p_tilde_hat))
  #beta_hat = solve(t(Z)%*%Z)%*%t(Z)%*%y

  #beta_hat
  #beta_hat/sum(beta_hat)

  Q = 0
  Sigma=0
  deltas = c()
  for(i in 1:length(y)){ 
    ag = Zu[i,]%*%t(Zu[i,])-diag(Zuu[i,])
    Q = Q + ag
    Delta = (ag%*%p_tilde_hat-y[i]*Zu[i,])
    Sigma = Sigma + Delta%*%t(Delta)
  }
  Q = Q/G
  Sigma = Sigma/G

  J = matrix(nrow = K,ncol=K)
  for(i in 1:K){
    for(j in 1:K){
      if(i==j){
        J[i,j] = sum(p_tilde_hat)-p_tilde_hat[i]
      }else{
        J[i,j] = -p_tilde_hat[i]
      }
    }
  }

  J = J/sum(p_tilde_hat)^2

  asyV = ((J)%*%solve(Q)%*%Sigma%*%solve(Q)%*%t(J))/G

  p_hat = p_tilde_hat/sum(p_tilde_hat)
  p_est = rbind(p_est,c(p_hat))
  
  p_tilde_var = rbind(p_tilde_var,diag(solve(Q)%*%Sigma%*%solve(Q)))
  
  p_var = rbind(p_var,diag(asyV))

}

ci_l = p_est - qnorm(0.975)*sqrt(p_var)
ci_r = p_est + qnorm(0.975)*sqrt(p_var)

#coverage

coverage = ((rep(1,100)%*%t(bulk_p))>=ci_l) & ((rep(1,100)%*%t(bulk_p))<=ci_r)
apply(coverage,2,mean)
```






how about change $beta$ to be $(0.4,0.3,0.2,0.1)$?




```{r}
G = nrow(Theta)
K = ncol(Theta)

set.seed(12345)
# bulk data library size: 300*number of genes.
# bulk_ls = rpois(1,300*nrow(Theta))
# total number of cells in bulk data
bulk_ncell = 400
# cell proportions
bulk_beta = c(4,3,2,1)
bulk_beta = bulk_beta/sum(bulk_beta)
bulk_p = (bulk_beta * (S))/sum(bulk_beta * (S))
# bulk data gene relative expression.
bulk_X = bulk_ncell*Theta%*%diag(S)%*%bulk_beta
bulk_theta = bulk_X/sum(bulk_X)

ref_ncell = 100
ref_X = Theta%*%diag(S)*ref_ncell

w = rep(1,G)


p_tilde_est = c()
p_tilde_var = c()
p_est = c()
p_var = c()

nrep = 100
for(rep in 1:nrep){
  if(rep%%200==0){print(rep)}
  bulk_ls = rpois(1,300*nrow(Theta))
  Z = matrix(rpois(G*K,ref_X),ncol=K)
  y = rpois(G,bulk_ls*bulk_theta)

  #W = diag(w)
  #beta_hat = solve(t(Z)%*%W%*%Z - diag(colSums(W%*%Z)))%*%t(Z)%*%W%*%y

  U = diag(colSums(Z))
  Zu = Z%*%(solve(U))
  Zuu = Zu%*%solve(U)
  p_tilde_hat = solve(t(Zu)%*%Zu - diag(colSums(Zuu)))%*%t(Zu)%*%y
  p_tilde_est = rbind(p_tilde_est,c(p_tilde_hat))
  #beta_hat = solve(t(Z)%*%Z)%*%t(Z)%*%y

  #beta_hat
  #beta_hat/sum(beta_hat)

  Q = 0
  Sigma=0
  deltas = c()
  for(i in 1:length(y)){ 
    ag = Zu[i,]%*%t(Zu[i,])-diag(Zuu[i,])
    Q = Q + ag
    Delta = (ag%*%p_tilde_hat-y[i]*Zu[i,])
    Sigma = Sigma + Delta%*%t(Delta)
  }
  Q = Q/G
  Sigma = Sigma/G

  J = matrix(nrow = K,ncol=K)
  for(i in 1:K){
    for(j in 1:K){
      if(i==j){
        J[i,j] = sum(p_tilde_hat)-p_tilde_hat[i]
      }else{
        J[i,j] = -p_tilde_hat[i]
      }
    }
  }

  J = J/sum(p_tilde_hat)^2

  asyV = ((J)%*%solve(Q)%*%Sigma%*%solve(Q)%*%t(J))/G

  p_hat = p_tilde_hat/sum(p_tilde_hat)
  p_est = rbind(p_est,c(p_hat))
  
  p_tilde_var = rbind(p_tilde_var,diag(solve(Q)%*%Sigma%*%solve(Q)))
  
  p_var = rbind(p_var,diag(asyV))

}

ci_l = p_est - qnorm(0.975)*sqrt(p_var)
ci_r = p_est + qnorm(0.975)*sqrt(p_var)

#coverage

coverage = ((rep(1,100)%*%t(bulk_p))>=ci_l) & ((rep(1,100)%*%t(bulk_p))<=ci_r)
apply(coverage,2,mean)
```




















































```{r,eval=FALSE,include=FALSE}
# Another dataset: 
XinT2D.eset = readRDS('data/deconv/XinT2Deset.rds')
XinT2D.eset
table(XinT2D.eset$cellType)
# remove genes 
rm.idx = which(rowSums((exprs(XinT2D.eset))!=0)<5)
rr = rowSums((exprs(XinT2D.eset)))

rm.idx2 = which(rr>=quantile(rr[-rm.idx],0.95))
rm.idx = unique(c(rm.idx,rm.idx2))
Theta = c()
S=c()
for(i in 1:4){
  aa = rowSums(exprs(XinT2D.eset)[-rm.idx,which(XinT2D.eset$cellTypeID == i)])
  Theta = cbind(Theta,aa/sum(aa))
  S[i] = sum(aa)/length(which(XinT2D.eset$cellTypeID == i))
}

cor(Theta)

S=S/100

set.seed(12345)
# bulk data library size: 50*number of genes.
bulk_ls = 50*nrow(Theta)
# total number of cells in bulk data
bulk_ncell = 400
# cell proportions
bulk_beta = c(1,2,3,4)
bulk_beta = bulk_beta/sum(bulk_beta)
# bulk data gene relative expression.
bulk_X = bulk_ncell*Theta%*%diag(S)%*%bulk_beta
bulk_theta = bulk_X/sum(bulk_X)

ref_ncell = 100
ref_X = Theta%*%diag(S)*ref_ncell

w = rep(1,length(bulk_theta))


ci_l = c()
ci_r = c()
beta_est = c()

for(rep in 1:100){
  
  Z = matrix(rpois(prod(dim(ref_X)),ref_X),ncol=ncol(ref_X))
y = rpois(length(bulk_theta),bulk_ls*bulk_theta)

#W = diag(w)
#beta_hat = solve(t(Z)%*%W%*%Z - diag(colSums(W%*%Z)))%*%t(Z)%*%W%*%y

beta_hat = solve(t(Z)%*%Z - diag(colSums(Z)))%*%t(Z)%*%y

#beta_hat = solve(t(Z)%*%Z)%*%t(Z)%*%y

#beta_hat
#beta_hat/sum(beta_hat)

Q = 0
Sigma=0
for(i in 1:length(y)){
  ag = Z[i,]%*%t(Z[i,])-diag(Z[i,])
  Q = Q + ag
  Delta = (ag%*%beta_hat-y[i]*Z[i,])
  Sigma = Sigma + Delta%*%t(Delta)
}
Q = Q*2/length(y)
Sigma = Sigma*4/length(y)

K = ncol(Z)
J = matrix(nrow = ncol(Z),ncol=ncol(Z))
for(i in 1:K){
  for(j in 1:K){
    if(i==j){
      J[i,j] = sum(beta_hat)-beta_hat[i]
    }else{
      J[i,j] = -beta_hat[i]
    }
  }
}

asyV = J%*%solve(Q)%*%Sigma%*%solve(Q)%*%J

beta_est = rbind(beta_est,c(beta_hat))

ci_l = cbind(ci_l,beta_hat/sum(beta_hat)-2*sqrt(diag(asyV)/length(y)))
ci_r = cbind(ci_r,beta_hat/sum(beta_hat)+2*sqrt(diag(asyV)/length(y)))
}

for(i in 1:4){
  plot(ci_l[i,],type='l',ylim=range(c(ci_l[i,],ci_r[i,])))
  lines(ci_r[i,],col='grey80')
  lines(rep(i/10,100),lty = 2)
}

```


```{r,eval=FALSE,include=FALSE}

set.seed(12345)

library(deconvSeq)

data("data_celltypes_rnaseq") 

dge.celltypes = getdge(cnts.celltypes, design.rnaseq, ncpm.min=1, nsamp.min=4)

A = dge.celltypes$counts%*%(design.rnaseq/(outer(rep(1,64),c(14,14,14,22))))

w = c(0.1,0.4,0.3,0.2)

######

Ey = A%*%w
e = rnorm(500,0,0.1)

y = Ey + e

# 1. ols est

what_ols = solve(t(A)%*%A)%*%t(A)%*%y
what_ols

# 2. nnls est

library(nnls)
what_nnls = nnls(A,y)
what_nnls$x

#######

c=10

Ey = c*A%*%w

y = Ey + e

what_ols = solve(t(A)%*%A)%*%t(A)%*%y


what_nnls = nnls(A,y)$x

D =t(A)%*%A

d = t(A)%*%y

Amat = diag(4)

Amat = rbind(rep(1,4),Amat)

b0 = c(1,0,0,0,0)

#what_qp = solve.QP(D,d,t(Amat),b0,meq=1)

#what_glmnet = glmnet(A,y,lambda = 0,lower.limits = 0,upper.limits = 1)

#what_bvls = bvls(A,y,rep(0,4),rep(1,4))

########

y = c*(A%*%w+e)

what_ols = solve(t(A)%*%A)%*%t(A)%*%y

nnls(A,y)$x

```

