---
title: "Methylation"
author: "DongyueXie"
date: "2020-01-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r,include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

DNA Mehylation vs enhancer/promoter?

What is nucleosome positions. Combine single cell? how? why? sliding window? 

Another way to use single cell data would be to get a measurement for each cell (with coverage) at regions that were identified as peaks or nucleosomes and that way get a measurement of heterogeneity.??

## BPRMeth

Inputs are Binomial DNA methylation data, with postions $x$. Tasks are 1. smooth each binomial sequences; 2. perform clustering on multiple DNA methylation sequences.(how? each seq is of different length) 3. binomial data can be bernoulli in single cell case. 


## BSmooth


## Introduction

Try BPRMeth package.

```{r,eval=FALSE}
library(BPRMeth)
met_region = encode_met
plot((encode_met$met[[21]])[,1],(encode_met$met[[21]])[,3]/(encode_met$met[[21]])[,2])

expr <- encode_expr
```


