---
title: "pbmcdata"
author: "Dongyue Xie"
date: "2020-01-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction


* Reproduce the experimental data used in Van den Berge et al., 2018 for creating PBMC null datasets. 

* Data includes 2,638 samples and 13,713 genes.


---

## Steps

Same steps as in `https://github.com/statOmics/zinbwaveZinger/blob/master/realdata/createdata/createDataObject.Rmd`. 


Reading in data.

```{r init, message=FALSE, eval=F}
library(Seurat)
library(dplyr)
library(Matrix)
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "~/Downloads/pbmc3k_filtered_gene_bc_matrices")
pbmc <- CreateSeuratObject(counts = pbmc.data, min.cells = 3, min.features = 200, project = "10X_PBMC")
```



QC and selecting cells for further analysis

```{r, eval=F}
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
```

Normalizing the data

```{r normalize, results='hide', eval=F}
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
```

Detection of variable genes across the single cells

```{r var_genes, fig.height=7, fig.width=11, results='hide', eval=F}
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
```

Scaling the data and removing unwanted sources of variation

```{r regress, fig.height=7, fig.width=11, results='hide', eval=F}
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes, vars.to.regress = "percent.mt")
```

Perform linear dimensional reduction

```{r pca, eval=F}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
```

Cluster the cells

```{r cluster, fig.height=5, fig.width=7, eval=F}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
```


Run Non-linear dimensional reduction (tSNE)

```{r tsne, fig.height=5, fig.width=7, message=FALSE, warnings=FALSE, eval=F}
pbmc <- RunUMAP(pbmc, dims = 1:10)
```

```{r, eval=F}
new.cluster.ids <- c("Naive CD4 T", "Memory CD4 T", "CD14+ Mono", "B", "CD8 T", "FCGR3A+ Mono", 
    "NK", "DC", "Platelet")
names(new.cluster.ids) <- levels(pbmc)
pbmc <- RenameIdents(pbmc, new.cluster.ids)
#DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```


Create SE object

```{r, eval=F}
library(SingleCellExperiment)
# use raw data as input for zinbwave but keep only non filtered cells
# and most variable genes indentified by seurat
pbmc2 <- CreateSeuratObject(counts = pbmc.data, min.cells = 3, min.features = 200, project = "10X_PBMC")
pbmc2[["percent.mt"]] <- PercentageFeatureSet(pbmc2, pattern = "^MT-")
pbmc2 <- subset(pbmc2, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)

keepcells = colnames(pbmc@data)
counts = pbmc2@assays$RNA[, keepcells]
# zinbwave does not want dgTMatrix as input
counts = as.matrix(counts)
counts = counts[rowSums(counts) > 0, ]
#keepcells = as.integer(pbmc@ident) %in% 1:2
#counts = counts[, keepcells]
# coldata
clusters = as.integer(pbmc@active.ident)
#clusters = clusters[clusters %in% 1:2]
cData = data.frame(seurat = clusters)
rownames(cData) = colnames(counts)
# rowdata
rData = data.frame(seuratVarGenes = rownames(counts) %in% rownames(pbmc@assays$RNA))
rownames(rData) = rownames(counts)
# create sce object
core = SingleCellExperiment(assays = list(counts = counts),
                            colData = cData, rowData = rData)
unloadNamespace("Seurat")
saveRDS(core, file = '~/Downloads/pbmc.rds')
saveRDS(assay(pbmc2), file = '~/Downloads/pbmc_counts.rds')
```


Cluster ID	Markers	Cell Type
0	IL7R, CCR7	Naive CD4+ T
1	IL7R, S100A4	Memory CD4+
2	CD14, LYZ	CD14+ Mono
3	MS4A1	B
4	CD8A	CD8+ T
5	FCGR3A, MS4A7	FCGR3A+ Mono
6	GNLY, NKG7	NK
7	FCER1A, CST3	DC
8	PPBP	Platelet
