---
title: "scde Calibration"
author: "DongyueXie"
date: "2020-03-08"
output:
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  chunk_output_type: console
---

## Introduction

Check if methods control claimed FDR.

Binomial thinning might not work very well in this case: consider a gene with expression $(1,1,1,1,1,1,1,1,0,0,...,0)$ in group 1 and $(1,1,0,0,0,0,0,...,0)$ in group 2 then if the effect is very small, then $p_2$ is close to 1 and $p_2$ is close to 0. So the thinned gene expression is likely to be $(0,0,0,0,0,...,0)$ and $(1,1,0,0,0,0,0,...,0)$. Even if we created signals for this gene, data are not informative enough.

How to avoid? For now just filter out genes appearing in less than 30 cells. (Total 709 cells). 

For fixed non-null gene set and gene effects, repeat 30 times: randomly assign samples to one of two groups and thin the data accordingly. 

```{r,message=FALSE,warning=FALSE,eval=FALSE}
library(vicar)
library(sva)
library(cate)
library(seqgendiff)
library(qvalue)
load('data/scde/scCDT.RData')

#'@param Z count matrix, sample by features
#'@param x 1 for group 1, 0 for group 2
#'@param beta effect of fearures,  0 for null.
#'@return W, thinned matrix
bi_thin = function(Z,x,beta){

  n=nrow(Z)
  p=ncol(Z)
  
  # group index
  g1 = which(x==1)
  g2 = which(x==0)
  
  
  p2 = 1/(1+exp(beta))
  p1 = 1-p2
  P = matrix(nrow = n,ncol = p)
  P[g1,] = t(replicate(length(g1),p1))
  P[g2,] = t(replicate(length(g2),p2))
  
  W = matrix(rbinom(n*p,Z,P),nrow=n)
  
  W
}

quiet <- function(x) { 
  sink(tempfile()) 
  on.exit(sink()) 
  invisible(force(x)) 
} 

```

```{r,eval=FALSE}
Z = t(as.matrix(CDT))
rm.idx = which(colSums(Z!=0)<30)
Z = Z[,-rm.idx]
n = nrow(Z)
p = ncol(Z)

cuts = c(1e-04, 0.001, 0.01, 0.025, 0.05,0.1)

set.seed(12345)
nreps = 50
#fix gene sets and effects
beta = rnorm(p,0,1.5)
beta[sample(1:p,p*0.9)]=0
which_null = ifelse(beta==0,1,0)

result_fdr = list()
result_auc = c()


for(rep in 1:nreps){
  x = rbinom(n,1,0.5)
  
  #generate data
  W = bi_thin(Z,x,beta)
  W = log(W + 0.5)
  
  X = model.matrix(~x)

  # fit sva
  sva_sva = quiet(sva(t(W),mod = X, mod0 = X[, -2, drop = FALSE], n.sv = 3))
  
  # sva-limma
  X.sva = cbind(X, sva_sva$sv)
  lmout = limma::lmFit(object = t(Wn), design = X.sva)
  eout  = limma::eBayes(lmout)

  svaout           <- list()
  svaout$betahat   <- lmout$coefficients[, 2]
  svaout$sebetahat <- lmout$stdev.unscaled[, 2] * sqrt(eout$s2.post)
  svaout$pvalues   <- eout$p.value[, 2]

  # ash
  sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)

  sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)

  #mouthwash
  
  mout1 = mouthwash(W,X,k=3,cov_of_interest = 2,
                   include_intercept = FALSE,verbose = FALSE,sprop = 1,scale_var = FALSE)
  
  mout0 = mouthwash(W,X,k=3,cov_of_interest = 2,
                   include_intercept = FALSE,verbose = FALSE,sprop = 0)
  
  
  
  #auc
  roc_out <- list(
  pROC::roc(response = which_null, predictor = c(mout$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(mout1$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(svaout$pvalues)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash0$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash1$result$lfsr)))
  

 auc_vec <- sapply(roc_out, FUN = function(x) { x$auc })
 

 result_auc = rbind(result_auc,auc_vec)
 name_vec <- c("MOUTHWASH0","MOUTHWASH1","SVA-limma","SVA-limma-ash0","SVA-limma-ash1")
 colnames(result_auc) <- name_vec
 
  # q value
  sva_qvalue = qvalue(svaout$pvalues)

  eFDR_sva = c()
  eFDR_sva_ash0 = c()
  eFDR_sva_ash1 = c()
  eFDR_mout1 = c()
  eFDR_mout0 = c()
  for(i in 1:length(cuts)){
    rej = which(sva_qvalue$qvalues<=cuts[i])
    eFDR_sva[i] = sum(rej%in%which(which_null==1))/length(rej)
  
    rej = which(mout1$result$qvalue<=cuts[i])
    eFDR_mout1[i] = sum(rej%in%which(which_null==1))/length(rej)
    
     rej = which(mout0$result$qvalue<=cuts[i])
    eFDR_mout0[i] = sum(rej%in%which(which_null==1))/length(rej)
  
    rej = which(sva_limma_ash1$result$qvalue<=cuts[i])
    eFDR_sva_ash1[i] = sum(rej%in%which(which_null==1))/length(rej)
    
    rej = which(sva_limma_ash0$result$qvalue<=cuts[i])
    eFDR_sva_ash0[i] = sum(rej%in%which(which_null==1))/length(rej)
    
  }
  
  result_fdr[[rep]] = rbind(eFDR_mout0,eFDR_mout1,eFDR_sva,eFDR_sva_ash0,eFDR_sva_ash1)
  save(result_fdr,file='data/scde/calibration_fdr.RData')
  save(result_auc,file='data/scde/calibration_auc.RData')
}
```


```{r}
load("~/misc/data/scde/calibration_auc3.RData")
load("~/misc/data/scde/calibration_fdr3.RData")

name_vec = c('mout0','mout1','sva','sva-ash0','sva-ash1')
colnames(result_auc) = name_vec
boxplot(result_auc,ylab = 'AUC')

cut1 = c()
cut2 = c()
cut3 = c()
cut4 = c()
for(i in 1:length(result_fdr)){
  cut1 = rbind(cut1,result_fdr[[i]][,3])
  cut2 = rbind(cut2,result_fdr[[i]][,4])
  cut3 = rbind(cut3,result_fdr[[i]][,5])
  cut4 = rbind(cut4,result_fdr[[i]][,6])
}

colnames(cut1) = name_vec
colnames(cut2) = name_vec
colnames(cut3) = name_vec
colnames(cut4) = name_vec
boxplot(cut1,ylab = 'FDP',ylim=c(0,0.03),main='fdr level 0.01')
abline(h=0.01,lty=3,col=4)
boxplot(cut2,ylab = 'FDP',ylim=c(0,0.055),main='fdr level 0.025')
abline(h=0.025,lty=3,col=4)
boxplot(cut3,ylab = 'FDP',ylim=c(0,0.1),main='fdr level 0.05')
abline(h=0.05,lty=3,col=4)
boxplot(cut4,ylab = 'FDP',ylim=c(0,0.2),main='fdr level 0.1')
abline(h=0.1,lty=3,col=4)

```






