---
title: "scde Check Alpha 0/1 on real data"
author: "DongyueXie"
date: "2020-03-19"
output:
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  chunk_output_type: console
---
```{r,include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE)
```



Compare the loglikelihood from sva-limma-ash, setting $\alpha=0$ and $\alpha=1$, on real data.

## Dataset 1: PBMC

### Summary

Even if we do not take log, $\alpha=1$ still give larger loglikelihood.

```{r}
library(vicar)
library(sva)
library(cate)
library(seqgendiff)
library(edgeR)
library(MASS)
load('data/scde/scCD4.RData')
load('data/scde/scCD8.RData')
load('data/scde/scCD14.RData')
load('data/scde/scMB.RData')

mglmfit = function(Y,X){
  G = nrow(Y)
  coefs = c()
  ses = c()
  for(i in 1:G){
    fit = glm(y~.,data.frame(y=Y[i,],x=X[,-1]),family = 'poisson')
    coefs = rbind(coefs,fit$coefficients)
    ses = rbind(ses,summary(fit)$coefficients[,2])
  }
  return(list(coefficients=coefs,se = ses))
}

```

CD4, CD8, CD14 and B cells. Look at number of genes and cells.

```{r}
dim(CD4)
dim(CD8)
dim(CD14)
dim(MB)
```


### CD4 vs MB cells.


```{r}
Y = as.matrix(cbind(CD4,MB))

group_idx = c(rep(1,dim(CD4)[2]),rep(0,dim(MB)[2]))

# remove genes appearing in less than 10 cells

Y = Y[-which(rowSums(Y!=0)<10),]

X = model.matrix(~group_idx)

real_study = function(Y,X,run_sva_eb=TRUE,n.sv=NULL){
  
  if(run_sva_eb){
    sva_sva = sva((Y),mod=X,mod0=X[,1],n.sv = n.sv)
    X.sva <- cbind(X, sva_sva$sv)
    lmout <- limma::lmFit(object = (Y), design = X.sva)
    eout  <- limma::eBayes(lmout)
    svaout           <- list()
    svaout$betahat   <- lmout$coefficients[, 2]
    svaout$sebetahat <- lmout$stdev.unscaled[, 2] * sqrt(eout$s2.post)
    svaout$pvalues   <- eout$p.value[, 2]
    sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)
    sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)
    print(paste('sva-limma-ash0 loglik: ',round(sva_limma_ash0$loglik,2),';',
            'sva-limma-ash1 loglik: ',round(sva_limma_ash1$loglik)))
  }else{
    lmout <- limma::lmFit(object = (Y), design = X)
    out = list()
    out$betahat   <- lmout$coefficients[, 2]
    out$sebetahat <- lmout$stdev.unscaled[, 2] * lmout$sigma
    lm_ash0 = ashr::ash(out$betahat,out$sebetahat,alpha=0)
    lm_ash1 = ashr::ash(out$betahat,out$sebetahat,alpha=1)
    print(paste('lm-ash0 loglik: ',round(lm_ash0$loglik,2),';',
            'lm-ash1 loglik: ',round(lm_ash1$loglik)))
  }
}

real_study(Y,X,3)
real_study(Y,X,run_sva_eb = FALSE)
```


Take log1p of Y

```{r}
real_study(log(Y+0.5),X,14)
real_study(log(Y+0.5),X,run_sva_eb = FALSE)
```

Voom-lm-ash

```{r}
d = DGEList(Y)
d = calcNormFactors(d)
y = voom(d, design=X, plot = FALSE)
real_study(y,X,run_sva_eb = FALSE)
```

glm-ash

```{r}
fit = mglmfit(Y,X)
glm_ash0 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=0)
glm_ash1 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=1)
print(paste('glm-ash0 loglik: ',round(glm_ash0$loglik,2),';',
            'glm-ash1 loglik: ',round(glm_ash1$loglik)))
```

### CD8 vs CD14 cells

```{r}
Y = as.matrix(cbind(CD8,CD14))

group_idx = c(rep(1,dim(CD8)[2]),rep(0,dim(CD14)[2]))

# remove genes appearing in less than 10 cells

Y = Y[-which(rowSums(Y!=0)<10),]

X = model.matrix(~group_idx)
real_study(Y,X,3)
real_study(Y,X,run_sva_eb = FALSE)

```


Take log1p of Y

```{r}
real_study(log(Y+0.5),X,27)
real_study(log(Y+0.5),X,run_sva_eb = FALSE)
```

Voom-lm-ash

```{r}
d = DGEList(Y)
d = calcNormFactors(d)
y = voom(d, design=X, plot = FALSE)
real_study(y,X,run_sva_eb = FALSE)
```

glm-ash

```{r}
fit = mglmfit(Y,X)
glm_ash0 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=0)
glm_ash1 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=1)
print(paste('glm-ash0 loglik: ',round(glm_ash0$loglik,2),';',
            'glm-ash1 loglik: ',round(glm_ash1$loglik)))
```

## Dataset 2: Jaitin et al.	2014

Taken from [here](https://github.com/jhsiao999/singleCellRNASeqMouseJaitinSpleen)

```{r}
load('data/scde/MouseJaitinSpleen.rda')
suppressPackageStartupMessages(library(SummarizedExperiment))
suppressPackageStartupMessages(library(MultiAssayExperiment))
mat = exprs(MouseJaitinSpleen)
idx = which(MouseJaitinSpleen$ERCC_dilution=='2.50E-05')
cell_type = MouseJaitinSpleen$group_name[idx]
table(cell_type)
mat = mat[,idx]
```

### B cell and NK_cell

```{r}
Y = as.matrix(cbind(mat[,which(cell_type=='B cell')],mat[,which(cell_type=='NK_cell')]))

group_idx = c(rep(1,dim(mat[,which(cell_type=='B cell')])[2]),
              rep(0,dim(mat[,which(cell_type=='NK_cell')])[2]))

# remove genes appearing in less than 10 cells

Y = Y[-which(rowSums(Y!=0)<10),]
rm.idx = which(colSums(Y)==0)
Y = Y[,-rm.idx]
group_idx = group_idx[-rm.idx]


X = model.matrix(~group_idx)
real_study(Y,X)
real_study(Y,X,run_sva_eb = FALSE)
```



```{r,eval=FALSE}

# How about just run linear model and then ash?

X = model.matrix(~group_idx)

lmout <- limma::lmFit(object = (Y), design = X)

svaout           <- list()
svaout$betahat   <- lmout$coefficients[, 2]
svaout$sebetahat <- lmout$stdev.unscaled[, 2] * lmout$sigma
svaout$pvalues   <- eout$p.value[, 2]

# sva-limma-ash-alpha 0

sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)
sva_limma_ash0$loglik

# sva-limma-ash-alpha 1

sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)
sva_limma_ash1$loglik
```

Take log1p of y

```{r}
real_study(log(Y+0.5),X)
real_study(log(Y+0.5),X,run_sva_eb = FALSE)
```

Voom-lm-ash

```{r}
d = DGEList(Y)
d = calcNormFactors(d)
y = voom(d, design=X, plot = FALSE)
real_study(y,X,run_sva_eb = FALSE)
```

glm-ash

```{r}
fit = mglmfit(Y,X)
glm_ash0 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=0)
glm_ash1 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=1)
print(paste('glm-ash0 loglik: ',round(glm_ash0$loglik,2),';',
            'glm-ash1 loglik: ',round(glm_ash1$loglik)))
```

### CD8-pDC  and splenocyte 

```{r}
Y = as.matrix(cbind(mat[,which(cell_type=='CD8-pDC')],mat[,which(cell_type=='splenocyte')]))

group_idx = c(rep(1,dim(mat[,which(cell_type=='CD8-pDC')])[2]),
              rep(0,dim(mat[,which(cell_type=='splenocyte')])[2]))

# remove genes appearing in less than 10 cells

Y = Y[-which(rowSums(Y!=0)<10),]
rm.idx = which(colSums(Y)==0)
Y = Y[,-rm.idx]
group_idx = group_idx[-rm.idx]

X = model.matrix(~group_idx)
real_study(Y,X)
real_study(Y,X,run_sva_eb = FALSE)

```


Take log1p of y

```{r}
real_study(log(Y+0.5),X)
real_study(log(Y+0.5),X,run_sva_eb = FALSE)
```

Voom-lm-ash

```{r}
d = DGEList(Y)
d = calcNormFactors(d)
y = voom(d, design=X, plot = FALSE)
real_study(y,X,run_sva_eb = FALSE)
```

glm-ash

```{r}
fit = mglmfit(Y,X)
glm_ash0 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=0)
glm_ash1 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=1)
print(paste('glm-ash0 loglik: ',round(glm_ash0$loglik,2),';',
            'glm-ash1 loglik: ',round(glm_ash1$loglik)))
```

## Dataset3: Grun et al 2016

Taken [here](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE81076)

```{r}
datax=readRDS("data/scde/GSE81076-GPL18573.rds")
datax_gene = experiments(datax)[["gene"]]
mat = (assays(datax_gene)[["count"]])
cell63_idx = 1:96
TGFBR3_idx = 673:768
Y = cbind(mat[,cell63_idx],mat[,TGFBR3_idx])
group_idx = c(rep(1,length(cell63_idx)),
              rep(0,length(TGFBR3_idx)))
Y = Y[-which(rowSums(Y!=0)<10),]

X = model.matrix(~group_idx)
real_study(Y,X)
real_study(Y,X,run_sva_eb = FALSE)
```

Take log1p of y

```{r}
real_study(log(Y+0.5),X)
real_study(log(Y+0.5),X,run_sva_eb = FALSE)
```

Voom-lm-ash

```{r}
d = DGEList(Y)
d = calcNormFactors(d)
y = voom(d, design=X, plot = FALSE)
real_study(y,X,run_sva_eb = FALSE)
```

glm-ash

```{r}
fit = mglmfit(Y,X)
glm_ash0 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=0)
glm_ash1 = ashr::ash(fit$coefficients[,2],fit$se[,2],alpha=1)
print(paste('glm-ash0 loglik: ',round(glm_ash0$loglik,2),';',
            'glm-ash1 loglik: ',round(glm_ash1$loglik)))
```

## Also compare log-limma-ash, voom-limma-ash and glm-ash on simulated and real data.
