---
title: "scde - check Mouthwash"
author: "DongyueXie"
date: "2020-02-28"
output:
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  chunk_output_type: console
---

## Introduction

Investigate why MOUTHWASH did not perform well in my previous [analysis](https://dongyuexie.github.io/misc/scdePBMC2.html).

Matthew suggested to look at the $\alpha$ value in `ash`. 

For each gene $j=1,2,...,G$, $\hat\beta_j$, estimated effect size and $\hat s_j$, standard error of the estimator are avaialbe from a statistical procedure like linear modelling. The t-statistics for testing $H_0: \beta_j = 0$  is $\frac{\hat\beta_j}{\hat s_j}$. 

`ash` prior: $\frac{\beta_j}{\hat{s}_j^\alpha}|\hat s_j \sim g(\cdot)$, $\alpha\geq 0$. When $\alpha>0$, this prior suggets effects with larger s.e. tend to be larger. When $\alpha=1$, lfsr is monotonic in p-values so ash-lfsr should give the same significant ranking as the one using p-values. 

A [note](binomialthinning.html) on why the simulated data shows effects with larger s.e. tend to be larger.

Compare the following methods:

`MOUTHWASH0`: Mouthwash, sprop=0 (corresonds to alpha=0 in ash), output lfsr

`MOUTHWASH1`: Mouthwash, sprop=1 (corresonds to alpha=1 in ash), output lfsr

`SVA-limma`: Run sva and limma , output p-value

`SVA-limma-ash0`: Run sva and limma, then run ash(alpha=0) with betahat and sehat, output lfsr

`SVA-limma-ash1`: Run sva and limma, then run ash(alpha=1) with betahat and sehat, output lfsr



```{r,message=FALSE,warning=FALSE}
library(vicar)
library(sva)
library(cate)
library(seqgendiff)
load('data/scde/scCDT.RData')
```

## exp signal

Add exponential distributed signal with rate=0.2.

### Filter out genes with 0 observation

```{r}
set.seed(12345)
CDT = as.matrix(CDT)
thinout = thin_2group(round(CDT),0.9,signal_fun = stats::rexp,signal_params = list(rate=0.2))
num_sv=3
Y = t(thinout$mat)
remove.idx = which(colSums(Y)==0)
eps=0.5
Y = log(Y[,-remove.idx]+eps)
X = model.matrix(~thinout$designmat)
which_null = c(1*(abs(thinout$coef[-remove.idx]) < 10^-6))
load('data/scde/mout_exp2_alpha0_0.RData')
load('data/scde/mout_exp2_alpha1_0.RData')
load('data/scde/sva_sva_exp2_0.RData')
X.sva <- cbind(X, sva_sva$sv)
lmout <- limma::lmFit(object = t(Y), design = X.sva)
eout  <- limma::eBayes(lmout)
svaout           <- list()
svaout$betahat   <- lmout$coefficients[, 2]
svaout$sebetahat <- lmout$stdev.unscaled[, 2] * sqrt(eout$s2.post)
svaout$pvalues   <- eout$p.value[, 2]

# sva-limma-ash-alpha 0

sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)
#sva_limma_ash0$loglik

# sva-limma-ash-alpha 1

sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)
#sva_limma_ash1$loglik

library(ggplot2)
################
roc_out <- list(
  pROC::roc(response = which_null, predictor = c(mout$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(mout1$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(svaout$pvalues)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash0$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash1$result$lfsr)))
name_vec <- c("MOUTHWASH0","MOUTHWASH1","SVA-limma","SVA-limma-ash0","SVA-limma-ash1")
names(roc_out) <- name_vec

sout <- lapply(roc_out, function(x) { data.frame(TPR = x$sensitivities, FPR = 1 - x$specificities)})
for (index in 1:length(sout)) {
  sout[[index]]$Method <- name_vec[index]
}
longdat <- do.call(rbind, sout)

shortdat <- dplyr::filter(longdat, Method == "MOUTHWASH0" | Method == "MOUTHWASH1" | 
                            Method == "SVA-limma" | Method == "SVA-limma-ash0" | Method == "SVA-limma-ash1")
ggplot(data = shortdat, mapping = aes(x = FPR, y = TPR, col = Method)) +
  geom_path() + theme_bw() + ggtitle("ROC Curves")

auc_vec <- sapply(roc_out, FUN = function(x) { x$auc })
knitr::kable(sort(auc_vec, decreasing = TRUE), col.names = "AUC", digits = 3)
```

### Filter out genes with less than 10 observation


```{r}
set.seed(12345)
thinout = thin_2group(round(CDT),0.9,signal_fun = stats::rexp,signal_params = list(rate=0.2))
num_sv=3
Y = t(thinout$mat)
remove.idx = which(colSums(Y>0)<=10)
eps=0.5
Y = log(Y[,-remove.idx]+eps)
X = model.matrix(~thinout$designmat)
which_null = c(1*(abs(thinout$coef[-remove.idx]) < 10^-6))
load('data/scde/mout_exp2_alpha0.RData')
load('data/scde/mout_exp2_alpha1.RData')
load('data/scde/sva_sva_exp2.RData')
X.sva <- cbind(X, sva_sva$sv)
lmout <- limma::lmFit(object = t(Y), design = X.sva)
eout  <- limma::eBayes(lmout)
svaout           <- list()
svaout$betahat   <- lmout$coefficients[, 2]
svaout$sebetahat <- lmout$stdev.unscaled[, 2] * sqrt(eout$s2.post)
svaout$pvalues   <- eout$p.value[, 2]

# sva-limma-ash-alpha 0

sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)
#sva_limma_ash0$loglik

# sva-limma-ash-alpha 1

sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)
#sva_limma_ash1$loglik

library(ggplot2)
################
roc_out <- list(
  pROC::roc(response = which_null, predictor = c(mout$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(mout1$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(svaout$pvalues)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash0$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash1$result$lfsr)))
name_vec <- c("MOUTHWASH0","MOUTHWASH1","SVA-limma","SVA-limma-ash0","SVA-limma-ash1")
names(roc_out) <- name_vec

sout <- lapply(roc_out, function(x) { data.frame(TPR = x$sensitivities, FPR = 1 - x$specificities)})
for (index in 1:length(sout)) {
  sout[[index]]$Method <- name_vec[index]
}
longdat <- do.call(rbind, sout)

shortdat <- dplyr::filter(longdat, Method == "MOUTHWASH0" | Method == "MOUTHWASH1" | 
                            Method == "SVA-limma" | Method == "SVA-limma-ash0" | Method == "SVA-limma-ash1")
ggplot(data = shortdat, mapping = aes(x = FPR, y = TPR, col = Method)) +
  geom_path() + theme_bw() + ggtitle("ROC Curves")

auc_vec <- sapply(roc_out, FUN = function(x) { x$auc })
knitr::kable(sort(auc_vec, decreasing = TRUE), col.names = "AUC", digits = 3)
```



## Normal signal with sd=0.8

Filter out genes with less than 10 observation

```{r}
set.seed(12345)
thinout = thin_2group(round(CDT),0.9,signal_fun = stats::rnorm,signal_params = list(mean=0,sd=0.8))
num_sv=3
Y = t(thinout$mat)
remove.idx = which(colSums(Y>0)<=10)
eps=0.5
Y = log(Y[,-remove.idx]+eps)
X = model.matrix(~thinout$designmat)
which_null = c(1*(abs(thinout$coef[-remove.idx]) < 10^-6))
load('data/scde/mout_normal8_alpha0.RData')
load('data/scde/mout_normal8_alpha1.RData')
load('data/scde/sva_sva_normal8.RData')
X.sva <- cbind(X, sva_sva$sv)
lmout <- limma::lmFit(object = t(Y), design = X.sva)
eout  <- limma::eBayes(lmout)
svaout           <- list()
svaout$betahat   <- lmout$coefficients[, 2]
svaout$sebetahat <- lmout$stdev.unscaled[, 2] * sqrt(eout$s2.post)
svaout$pvalues   <- eout$p.value[, 2]

# sva-limma-ash-alpha 0

sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)
#sva_limma_ash0$loglik

# sva-limma-ash-alpha 1

sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)
#sva_limma_ash1$loglik

library(ggplot2)
################
roc_out <- list(
  pROC::roc(response = which_null, predictor = c(mout$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(mout1$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(svaout$pvalues)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash0$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash1$result$lfsr)))
name_vec <- c("MOUTHWASH0","MOUTHWASH1","SVA-limma","SVA-limma-ash0","SVA-limma-ash1")
names(roc_out) <- name_vec

sout <- lapply(roc_out, function(x) { data.frame(TPR = x$sensitivities, FPR = 1 - x$specificities)})
for (index in 1:length(sout)) {
  sout[[index]]$Method <- name_vec[index]
}
longdat <- do.call(rbind, sout)

shortdat <- dplyr::filter(longdat, Method == "MOUTHWASH0" | Method == "MOUTHWASH1" | 
                            Method == "SVA-limma" | Method == "SVA-limma-ash0" | Method == "SVA-limma-ash1")
ggplot(data = shortdat, mapping = aes(x = FPR, y = TPR, col = Method)) +
  geom_path() + theme_bw() + ggtitle("ROC Curves")

auc_vec <- sapply(roc_out, FUN = function(x) { x$auc })
knitr::kable(sort(auc_vec, decreasing = TRUE), col.names = "AUC", digits = 3)
```


## Normal signal with sd=1.5

Filter out genes with less than 10 observation


```{r}
set.seed(12345)
thinout = thin_2group(round(CDT),0.9,signal_fun = stats::rnorm,signal_params = list(mean=0,sd=1.5))
num_sv=3
Y = t(thinout$mat)
remove.idx = which(colSums(Y>0)<=10)
eps=0.5
Y = log(Y[,-remove.idx]+eps)
X = model.matrix(~thinout$designmat)
which_null = c(1*(abs(thinout$coef[-remove.idx]) < 10^-6))
load('data/scde/mout_normal15_alpha0.RData')
load('data/scde/mout_normal15_alpha1.RData')
load('data/scde/sva_sva_normal15.RData')
X.sva <- cbind(X, sva_sva$sv)
lmout <- limma::lmFit(object = t(Y), design = X.sva)
eout  <- limma::eBayes(lmout)
svaout           <- list()
svaout$betahat   <- lmout$coefficients[, 2]
svaout$sebetahat <- lmout$stdev.unscaled[, 2] * sqrt(eout$s2.post)
svaout$pvalues   <- eout$p.value[, 2]

# sva-limma-ash-alpha 0

sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)
#sva_limma_ash0$loglik

# sva-limma-ash-alpha 1

sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)
#sva_limma_ash1$loglik

library(ggplot2)
################
roc_out <- list(
  pROC::roc(response = which_null, predictor = c(mout$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(mout1$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(svaout$pvalues)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash0$result$lfsr)),
  pROC::roc(response = which_null, predictor = c(sva_limma_ash1$result$lfsr)))
name_vec <- c("MOUTHWASH0","MOUTHWASH1","SVA-limma","SVA-limma-ash0","SVA-limma-ash1")
names(roc_out) <- name_vec

sout <- lapply(roc_out, function(x) { data.frame(TPR = x$sensitivities, FPR = 1 - x$specificities)})
for (index in 1:length(sout)) {
  sout[[index]]$Method <- name_vec[index]
}
longdat <- do.call(rbind, sout)

shortdat <- dplyr::filter(longdat, Method == "MOUTHWASH0" | Method == "MOUTHWASH1" | 
                            Method == "SVA-limma" | Method == "SVA-limma-ash0" | Method == "SVA-limma-ash1")
ggplot(data = shortdat, mapping = aes(x = FPR, y = TPR, col = Method)) +
  geom_path() + theme_bw() + ggtitle("ROC Curves")

auc_vec <- sapply(roc_out, FUN = function(x) { x$auc })
knitr::kable(sort(auc_vec, decreasing = TRUE), col.names = "AUC", digits = 3)
```



```{r,eval=FALSE}
set.seed(12345)
CDT = as.matrix(CDT)

# add signals to null data, 90% are null, 10% signals.

thinout = thin_2group(round(CDT),0.9,signal_fun = stats::rexp,signal_params = list(rate=0.2))

#thinout = thin_2group(round(CDT),0.9,signal_fun = stats::rnorm,signal_params = list(mean=0,sd=1.5))
num_sv=3
Y = t(thinout$mat)

# for now just remove genes with no observations. 

#### could also try to remove genes with less than 10 non-zero appearances.###

remove.idx = which(colSums(Y>0)<=10)
eps=0.5
Y = log(Y[,-remove.idx]+eps)
X = model.matrix(~thinout$designmat)
which_null = c(1*(abs(thinout$coef[-remove.idx]) < 10^-6))

# directly apply ash to sva-limma with alpha=0 or 1 and compare results with mouthwash(alpha=0 or 1) and sva
# rank genes by p value and lfsr

# mouthwash alpha = 0
#mout = mouthwash(Y,X,k=num_sv,cov_of_interest = 2,include_intercept = FALSE,sprop = 0,scale_var = TRUE)
#save(mout,file = 'data/scde/mout_exp2_alpha0.RData')
load('data/mout_exp2_alpha0.RData')

# mouthwash alpha = 1
#mout1 = mouthwash(Y,X,k=num_sv,cov_of_interest = 2,include_intercept = FALSE,sprop = 1,scale_var = FALSE)
#save(mout1,file = 'data/scde/mout_exp2_alpha1.RData')
load('data/mout_high_alpha1.RData')

# sva-limma
#sva_sva = sva::sva(dat = t(Y), mod = X, mod0 = X[, -2, drop = FALSE], n.sv = num_sv)
#save(sva_sva,file = 'data/scde/sva_sva_exp2.RData')
load('data/sva_sva_high.RData')

X.sva <- cbind(X, sva_sva$sv)
lmout <- limma::lmFit(object = t(Y), design = X.sva)
eout  <- limma::eBayes(lmout)
svaout           <- list()
svaout$betahat   <- lmout$coefficients[, 2]
svaout$sebetahat <- lmout$stdev.unscaled[, 2] * sqrt(eout$s2.post)
svaout$pvalues   <- eout$p.value[, 2]

# sva-limma-ash-alpha 0

sva_limma_ash0 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=0)
sva_limma_ash0$loglik

# sva-limma-ash-alpha 1

sva_limma_ash1 = ashr::ash(svaout$betahat,svaout$sebetahat,alpha=1)
sva_limma_ash1$loglik
```

```{r,eval=FALSE}
library(plotrix)
plotCI(svaout$sebetahat,(svaout$betahat),ui = svaout$betahat + 2*svaout$sebetahat,li = svaout$betahat - 2*svaout$sebetahat,pch=20,sfrac = 0.005,xlab='standard error', ylab='estimated effects')
```


```{r,eval=FALSE}

## Rank genes

#Rank genes based on lfsr and p-values.

mout0_rank = order(mout$result$lfsr,decreasing = F)
mout1_rank = order(mout1$result$lfsr,decreasing = F)
svalimma_rank = order(svaout$pvalues,decreasing = F)
svalimma_ash0_rank = order(sva_limma_ash0$result$lfsr,decreasing = F)
svalimma_ash1_rank = order(sva_limma_ash1$result$lfsr,decreasing = F)

# Compare top 10 genes ranked by `sva-limma` and `sva-limma-ash1`.
ng=10
ranks = cbind(svalimma_rank[1:ng],svalimma_ash1_rank[1:ng])
colnames(ranks) = c('sva-limma','sva-limma-ash1')
rownames(ranks) = 1:ng
knitr::kable(ranks)

# The first 6 are different. How about the rest genes? From the plot below, the ranks are the same.
plot(svalimma_rank[-c(1:6)],svalimma_ash1_rank[-c(1:6)],xlab = 'gene ranks - sva_limma_p-value',ylab = 'gene ranks - sva_limma_ash1')

#Why the first 6 genes are different? Let's look at the p-values and lfsr. The first 6 genes have 0 lfsr due to numerical precision limit I guess. 

svaout$pvalues[svalimma_rank[1:6]]
sva_limma_ash1$result$lfsr[svalimma_ash1_rank[1:6]]
```


### Summary

1. Setting alpha=1 outperforms setting alpha=0. 
2. Methods favor filtering out genes with less than 10 observations 

## Calibration

Let's use the example of Normal signals with sd=1.5.

At a given FDR do they achieve what they claim.

```{r}
library(qvalue)
sva_qvalue = qvalue(svaout$pvalues)
cuts = c(1e-04, 0.001, 0.01, 0.025, 0.05,0.1)
eFDR_sva = c()
eFDR_sva_ash = c()
eFDR_mout = c()
for(i in 1:length(cuts)){
  rej = which(sva_qvalue$qvalues<=cuts[i])
  eFDR_sva[i] = sum(rej%in%which(which_null==1))/length(rej)
  
  rej = which(mout1$result$qvalue<=cuts[i])
  eFDR_mout[i] = sum(rej%in%which(which_null==1))/length(rej)
  
  rej = which(sva_limma_ash1$result$qvalue<=cuts[i])
  eFDR_sva_ash[i] = sum(rej%in%which(which_null==1))/length(rej)
}

knitr::kable(cbind(cuts,round(eFDR_sva,5),round(eFDR_sva_ash,5),round(eFDR_mout,5)),
             col.names = c('fdr_level','sva','sva_ash','mouthwash'))
```

